---
title: "Moderated Latent Variables Mediation Using Correct and Misspecified Models"
shorttitle: "Moderated Mediation With Latent Variables"

author:
  - name: Felipe Fontana Vieira
    corresponding: true
    orcid: 0009-0006-0949-6569
    email: felipe.fontanavieira@ugent.be
    affiliations:
      - name: Ghent University
        department: Department of Data Analysis, Faculty of Psychology and Educational Sciences
        address: Henri Dunantlaan 2
        city: Ghent
        country: Belgium
        postal-code: 9000
        id: UGent
    roles:
      - Conceptualization
      - Methodology
      - Software
      - Investigation
      - Data curation
      - Formal analysis
      - Visualization
      - Writing - original draft
      - Writing - review & editing
  - name: Yves Rosseel
    corresponding: false
    orcid: 0000-0002-4129-4477
    affiliations:
      - ref: UGent
    roles:
      - Conceptualization
      - Methodology
      - Software
      - Supervision
      - Writing - review & editing
      - Funding acquisition

author-note:
  status-changes:
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: "This study was preregistered."
    data-sharing: "The preregistration, data, code, and materials for this 
    manuscript can be found at <https://doi.org/10.5281/zenodo.20703258>."
    financial-support: "This project was funded by the Special Research Fund 
    (BOF) of Ghent University (BOF/24J/2023/144)."

abstract: |
  TODO: Abstract.

impact-statement: |
  TODO: Translational abstract.

keywords: [moderated mediation, structural equation modeling, latent interaction, measurement error, model misspecification]

language:
  title-impact-statement: "Translational Abstract"

word-count: true
floatsintext: true
numbered-lines: false
link-citations: false
mask: false

bibliography: "references.bib"

format:
  apaquarto-pdf:
    documentmode: man
    keep-tex: false
    include-in-header:
      text: |
        \raggedbottom
        \usepackage{pdflscape}

execute:
  echo: false
  warning: false
  message: false
---

Mediation analysis is a well-known procedure in the behavioral and social
sciences [@Hayes2018; @Preacher2007; @Pieters2017; @Charlton2021]. In its 
basic and standard form, the effect of an independent variable on an outcome is 
decomposed into a direct effect and an indirect effect transmitted through a 
mediator, with all variables observed. This basic single-mediator model has 
since been extended, for example by allowing the indirect effect to depend on 
a moderator and by adjusting for covariates [@Hayes2015; @Edwards2007; 
@Preacher2007; @Montoya2024].

A more fundamental concern arises from measurement. Many constructs in 
psychology are not directly observed but are measured indirectly through 
multiple items, that is, they are latent variables [@BollenHoyle2012;
@LedgerwoodShrout2011].
When such constructs are reduced to summary scores 
(e.g., a sum or mean of the items) and analyzed as though they were observed, 
measurement error is ignored. This error can bias the indirect effect and 
inflate its uncertainty [@HoyleKenny1999; @ColePreacher2014], and it is 
especially problematic for the mediator, which acts at once as a predictor and 
an outcome [@GonzalezMacKinnon2021]. To address this, SEM with latent variables
has been proposed [@cite], especially in the context of mediation [@cite]. 
It counteracts the attenuation by estimating structural
relationships between common factors, from which indicator-specific unique
variance has been partitioned [@Bollen1989], provided the measurement model is
correctly specified [@cite]. The problem is compounded under moderation, 
where the product term is less reliable than either of its constituents 
[@Bohrnstedt1978; @Busemeyer1983].

Yet, combining latent measurement with moderation yields a latent-variable 
moderated mediation model [@Hayes2015], which is not straightforward in terms of
estimation. A latent interaction involves the product of two latent 
variables, which is unobserved and nonnormal even when its constituents are 
normal [@KleinMoosbrugger2000]. A number of methods have been proposed to
estimate latent interactions [@AlginaMoulder2001; @Arminger1998;
@AsparouhovMuthen2021; @Boker2023; @Bollen1995; @CoxKelcey2021; @HolstBudtz2020;
@JoreskogYan1996; @KennyJudd1984; @KleinMoosbrugger2000; @KleinMuthen2007;
@Marsh2004; @MooijaartBentler2010; @MooijaartSatorra2012; @NgChan2020;
@RosseelLoh2024; @WallAmemiya2000; @WallAmemiya2001; @WallAmemiya2003]. 
Most adopt a system-wide (joint) estimation
strategy, in which the measurement and structural parameters are estimated
simultaneously, and include the most known methods: latent moderated
structural equations (LMS) and product-indicator (PI). Opposed to these 
are structural-after-measurement (SAM) approaches [@CoxKelcey2021; 
@HolstBudtz2020; @NgChan2020; @RosseelLoh2024],
in which the measurement model is estimated first, from the indicators alone, 
and the structural model is estimated afterward with the measurement parameters 
held fixed. Until recently, these abovementioned approaches were rarely applied 
in practice, held back by limited software, slow estimation, and error-prone 
implementations. This has changed with recent open-source implementations 
[@Slupphaug2025; @Rosseel2025; @HolstBudtz2020] and non-commercial software 
[@Keller2025].

In this context, a further consideration concerns interpretation. Mediation is 
a hypothesis about causal effects, in that the indirect effect is intended as 
the effect of the predictor on the outcome that is transmitted through the 
mediator. A causal interpretation requires identification assumptions that 
cannot be verified from the data alone [@Imai2010; @Rohrer2022; 
@VanderWeele2015]. With latent variables, further requirements apply 
[@Sterner2024; @VanderWeeleVansteelandt2022]. We focus on one that is central 
to interpretation: the estimated latent variables must validly represent the 
intended constructs [@GonzalezMacKinnon2021]. When they do not, the model is 
subject to interpretational confounding [@Burt1976], a discrepancy between the 
empirical meaning assigned to a latent variable through its estimated 
measurement parameters and the meaning intended by the researcher. 
In one-step SEM, where the measurement and structural parameters are estimated
simultaneously, a misspecification in the structural model can propagate to the
measurement model and alter the estimated latent variables. In moderated
mediation, such misspecifications are plausible (e.g., an omitted moderator 
effect on the outcome, an omitted moderated direct effect, a cross-loading, or 
a residual mediator-outcome covariance), and each can distort the constructs 
whose conditional indirect effect is of interest. SAM avoids this problem. 
Because the structural model
cannot influence the measurement model, structural misspecification does not 
affect the estimated latent variables, and interpretational confounding is
avoided, though this does not imply that the structure is correct or fixed. 
Local SAM (LSAM) is one variant of this approach, and it has recently been 
extended to interaction and quadratic effects among latent variables 
[@Rosseel2025], which makes it applicable to latent-variable moderated 
mediation.

To illustrate, for simplicity, we take a mediation model with three latent 
variables, X, M, and Y, each with three indicators, and obtain a population 
covariance matrix 
from a correctly specified version. We then alter it: the covariances between 
the X and M indicators are made unequal across the X indicators (3.0, 1.5, and 
1.5, with mean held at the implied value of 2.0), so that only their pattern, 
not the overall X-M strength, changes. We fit the model by one-step SEM and by 
SAM ($N = 200$).

```{r}
#| label: tbl-interp-confounding
#| tbl-cap: "Illustration for Interpretational Confounding."
library(lavaan)

model.pop <- '
  X =~ 1*x1 + 1*x2 + 1*x3
  Y =~ 1*y1 + 1*y2 + 1*y3
  M =~ 1*m1 + 1*m2 + 1*m3
  X ~~ 4*X
  M ~~ 3.0*M
  Y ~~ 2.76*Y
  x1 ~~ 1.7*x1; x2 ~~ 1.7*x2; x3 ~~ 1.7*x3
  y1 ~~ 1.7*y1; y2 ~~ 1.7*y2; y3 ~~ 1.7*y3
  m1 ~~ 1.7*m1; m2 ~~ 1.7*m2; m3 ~~ 1.7*m3
  M ~ 0.5*X
  Y ~ 0.1*X + 0.5*M
'
Sigma <- lavInspect(sem(model.pop), "implied")$cov

Sigma2 <- Sigma
Sigma2[c("m1", "m2", "m3"), "x1"] <- 3.0
Sigma2[c("m1", "m2", "m3"), "x2"] <- 1.5
Sigma2[c("m1", "m2", "m3"), "x3"] <- 1.5
Sigma2["x1", c("m1", "m2", "m3")] <- 3.0
Sigma2["x2", c("m1", "m2", "m3")] <- 1.5
Sigma2["x3", c("m1", "m2", "m3")] <- 1.5

model <- '
  X =~ x1 + x2 + x3
  Y =~ y1 + y2 + y3
  M =~ m1 + m2 + m3
  M ~ X
  Y ~ M + X
'
fit_sem <- sem(model, sample.cov = Sigma2, sample.cov.rescale = FALSE, 
               sample.nobs = 200L)
fit_sam <- sam(model, sample.cov = Sigma2, sample.cov.rescale = FALSE, 
               sample.nobs = 200L)

pull <- function(fit) {
  est <- lavInspect(fit, "est")
  L <- est$lambda
  B <- est$beta
  c(L["x1", "X"], L["x2", "X"], L["x3", "X"],
    B["M", "X"], B["Y", "M"], B["Y", "X"], B["M", "X"] * B["Y", "M"])
}

tab <- data.frame(
  Parameter = c("Loading of $x_1$ on X", "Loading of $x_2$ on X",
                "Loading of $x_3$ on X", "$M \\sim X$", "$Y \\sim M$",
                "$Y \\sim X$", "Indirect effect"),
  True = c(1, 1, 1, 0.5, 0.5, 0.1, 0.25),
  `One-step SEM` = pull(fit_sem),
  SAM = pull(fit_sam),
  check.names = FALSE
)
knitr::asis_output(paste0("\\begin{center}\n",
  knitr::kable(tab, digits = 2, booktabs = TRUE, escape = FALSE, 
               format = "latex"),
  "\n\\end{center}"))
```

As shown in @tbl-interp-confounding, one-step SEM yields a proper solution but
distorts the X loadings and biases the structural paths, with the direct effect 
of X on Y falling from 0.10 to 0.03. SAM estimates the measurement model from 
the indicators alone, leaving the loadings intact and recovering the true 
parameters.

### Present Research and Contributions

This study evaluates LSAM, alongside LMS and a double-mean-centered 
product-indicator approach, for estimating the index of moderated mediation.
It makes three contributions.

First, inference for the index has relied on the bootstrap in the limited 
prior work [@Hayes2015; @Tibbe2022; @FalkBiesanz2015; @Feng2020]. We instead 
use the Monte Carlo method and evaluate how well it is calibrated.

Second, these estimators, and LSAM and LMS in particular, have not been examined 
for moderated mediation under structural misspecification. We evaluate them 
under a battery of plausible misspecifications (an omitted $W \to Y$ effect, 
an omitted moderated direct effect, an omitted cross-loading, and an omitted 
residual covariance, each at two magnitudes), the omissions that induce 
interpretational confounding under one-step estimation.

Third, LSAM has not been evaluated in a mediation setting or for derived 
quantities such as the indirect effect $a\,b$ or the index $a_3 b$, which are 
nonlinear functions of the structural parameters whose behavior need not match 
that of their constituents. We assess all three estimators across realistic 
sample sizes, two reliability levels, and nonnormal latent predictors, on bias, 
root-mean-square error, coverage, and the Type I error and power of the index, 
and close with practical recommendations for estimator choice.

```{r}
#| include: false
# smoke-test for the knitr engine. add library() calls and data reads
# (e.g. readRDS("../results/summary.rds")) — plus the matching packages in
# gen_env.R's r_pkgs — as the manuscript needs them.
invisible(1 + 1)
```

# Simulation Design and Performance Evaluation

In this section we describe the design of the simulation study and the
measures used to evaluate performance, following recent recommendations for 
the conduct and reporting of simulation studies [@Morris2019; @Siepe2024; @Pawel2026].

### Population Model and Estimand

We study a moderated mediation model among four latent variables: an
exogenous predictor $X$, a moderator $W$, a mediator $M$, and an outcome $Y$.
The structural model is as follows

$$
M = a_1 X + a_2 W + a_3\,(X \times W) + \zeta_M, \qquad
Y = b\,M + c'\,X + \zeta_Y,
$$

The primary estimands are the index of moderated mediation (imm) [@Hayes2015],
i.e., $a_3 b$. The remaining
structural coefficients ($a_1$, $a_2$, $b$, $c'$) and the twelve freely estimated
factor loadings serve as secondary estimands. Because the index is a derived
quantity rather than a free parameter, its true value is computed as $a_3 b$
implied by the correctly specified model, and each estimator's recovery of that
value is assessed across conditions.

Population values were fixed at $a_1 = 0.4$, $a_2 = 0.3$, $b = 0.5$, and
$c' = 0.15$, with $\mathrm{cor}(X, W) = 0.3$. The interaction coefficient $a_3$
is a design factor (see below). These coefficients, together with the structural
disturbance variances, were chosen so that under the baseline condition (normal
exogenous predictors and a correctly specified model) each dependent latent
variable has $R^2 = 0.30$ at the interaction level $a_3 = 0.20$. Note that, 
because $a_3$ contributes to the variance of $M$, the explained variance is 
naturally higher at $a_3 = 0.40$ and lower at $a_3 = 0$. Each latent variable is 
measured by four congeneric indicators with loadings $1.0$, $0.8$, $0.7$, 
and $0.6$. @fig-model depicts the population
model, where the structural misspecifications studied below are drawn in grey.

```{tikz}
%| label: fig-model
%| fig-cap: "Population (Data-Generating) Model for the Simulation Study."
%| apa-note: "*Note*. $X$, $W$, $M$, and $Y$ are latent variables, each measured by four congeneric indicators; circles are latent variables and $X{\\times}W$ is the latent interaction. Black elements form the correctly specified model that every estimator fits ($a_1$, $a_2$, $a_3$, $b$, $c'$, and the $X$--$W$ covariance). Grey elements are the misspecifications present in the population but omitted from the fitted model: two in the structural part, an omitted moderator effect on the outcome ($c_2$) and an omitted moderated direct effect ($c_3$); and two in the measurement part, an omitted cross-loading of $X$ on $m_3$ (cl) and an omitted residual covariance between $m_3$ and $y_3$ (rcov). Residual variances and intercepts are not shown."
%| fig-align: center
%| fig-width: 9
%| fig-height: 6

\tikzset{
  latent/.style  ={circle,draw,minimum size=12mm},
  ind/.style     ={rectangle,draw,minimum height=6mm,minimum width=7mm,align=center},
  derived/.style ={circle,draw,minimum size=12mm},
  mis/.style     ={->,gray!70,thick,dashed},
  miscov/.style  ={<->,gray!70,thick,dashed},
  every node/.style={font=\small}
}

\begin{tikzpicture}

% X
\node[ind] (x1) at (0,  5.4) {$x_1$};
\node[ind] (x2) at (0,  4.6) {$x_2$};
\node[ind] (x3) at (0,  3.8) {$x_3$};
\node[ind] (x4) at (0,  3.0) {$x_4$};
\node[latent] (X) at (2.6, 4.2) {$X$};
\draw[->] (X.west) -- (x1.east);
\draw[->] (X.west) -- (x2.east);
\draw[->] (X.west) -- (x3.east);
\draw[->] (X.west) -- (x4.east);

% moderator W
\node[ind] (w1) at (0,  1.4) {$w_1$};
\node[ind] (w2) at (0,  0.6) {$w_2$};
\node[ind] (w3) at (0, -0.2) {$w_3$};
\node[ind] (w4) at (0, -1.0) {$w_4$};
\node[latent] (W) at (2.6, 0.2) {$W$};
\draw[->] (W.west) -- (w1.east);
\draw[->] (W.west) -- (w2.east);
\draw[->] (W.west) -- (w3.east);
\draw[->] (W.west) -- (w4.east);

% interaction
\node[derived] (XW) at (2.6, -3.2) {$X{\times}W$};

% X-W covariance (part of the fitted model)
\draw[<->] (X) to[bend right=42] (W);

% m
\node[latent] (M) at (7.0, 2.2) {$M$};
\node[ind] (m1) at (5.5, 5.2) {$m_1$};
\node[ind] (m2) at (6.5, 5.2) {$m_2$};
\node[ind] (m3) at (7.5, 5.2) {$m_3$};
\node[ind] (m4) at (8.5, 5.2) {$m_4$};
\draw[->] (M.north) -- (m1.south);
\draw[->] (M.north) -- (m2.south);
\draw[->] (M.north) -- (m3.south);
\draw[->] (M.north) -- (m4.south);

% outcome y
\node[latent] (Y) at (11.4, 2.2) {$Y$};
\node[ind] (y1) at (13.8, 3.4) {$y_1$};
\node[ind] (y2) at (13.8, 2.6) {$y_2$};
\node[ind] (y3) at (13.8, 1.8) {$y_3$};
\node[ind] (y4) at (13.8, 1.0) {$y_4$};
\draw[->] (Y.east) -- (y1.west);
\draw[->] (Y.east) -- (y2.west);
\draw[->] (Y.east) -- (y3.west);
\draw[->] (Y.east) -- (y4.west);

% correct struct paths
\draw[->] (X)  -- node[above,sloped,font=\footnotesize] {$a_1$} (M);
\draw[->] (W)  -- node[below,sloped,font=\footnotesize] {$a_2$} (M);
\draw[->] (XW) -- node[below,sloped,font=\footnotesize] {$a_3$} (M);
\draw[->] (M)  -- node[above,font=\footnotesize] {$b$} (Y);
\draw[->] (X)  to[bend left=22] node[above,font=\footnotesize] {$c'$} (Y);

% misspecifications (grey): c2/c3 structural, cl/rcov measurement
\draw[mis] (W)  to[bend right=24] node[below,sloped,font=\footnotesize,text=gray] {$c_2$} (Y);
\draw[mis] (XW) to[bend right=34] node[below,sloped,font=\footnotesize,text=gray] {$c_3$} (Y);
\draw[mis] (X) to[out=68,in=128] node[above,font=\footnotesize,text=gray,pos=0.62] {cl} (m3.north);
\draw[miscov] (m3.north) to[out=42,in=150] node[above,sloped,font=\footnotesize,text=gray,pos=0.5] {rcov} (y3.north west);

\end{tikzpicture}
```

### Simulation Conditions

We employed a fully factorial design crossing five factors: sample size
($n \in \{100, 200, 300, 500\}$), interaction magnitude
($a_3 \in \{0, 0.20, 0.40\}$), measurement reliability (low $= 0.5$,
high $= 0.7$), the distribution of the exogenous latent predictors (normal, 
uniform, $t$, and two right-skewed $\chi^2$ distributions), 
and the (mis)specification condition (nine levels; see
below). This yielded $4 \times 3 \times 2 \times 5 \times 9 = 1{,}080$ 
conditions, each replicated 1,000 times.

Sample sizes span the small-to-moderate range typical of applied moderated
mediation research, where interaction terms are estimated with limited power
[@Hayes2015; @Preacher2007]. The interaction magnitudes include a null condition
($a_3 = 0$) to assess Type I error and false-positive coverage of the index, and
two nonzero levels representing small and moderate latent moderation effects.
Reliability levels of $0.5$ and $0.7$ were chosen because measurement error is
amplified in product terms and is therefore especially consequential for latent
interaction estimation [@Bollen1989; @Brandt2014], and because lower reliability
is common in applied settings than the values typically examined in prior work.
Nonnormality of the latent predictors is both empirically common and methodologically central for latent interaction estimators, whose distributional assumptions differ across approaches [@KleinMoosbrugger2000; @KleinMuthen2007; @Cham2012]. The misspecification conditions follow the moderated-mediation framework for the
two structural omissions [@Preacher2007] and prior latent-interaction simulation
work for the two measurement omissions [@Brandt2020]. They were chosen as
interpretable threats, acting respectively as a measured mediator–outcome
confounder, a conflated moderated direct effect, exposure-correlated measurement
error, and an unmeasured residual confounder [@VanderWeele2015; @Edwards2007]. The
nine levels are detailed below.

### Estimators

We compared a two-step estimator against two system-wide estimators. The
two-step approach is LSAM [@RosseelLoh2024; @Rosseel2025]. In this case, the 
interacting predictors $X$ and $W$ shared a single measurement block, 
while $M$ and $Y$ formed their own blocks. The two
system-wide estimators are LMS [@KleinMoosbrugger2000] and UPI with double 
mean centering [@KelavaBrandt2009; @Marsh2004; @Lin2010].

### Data Generation

Appendix A describes the complete data-generating mechanism in detail. Briefly,
the latent variables $X$ and $W$ were drawn from a bivariate
distribution with $\operatorname{Cor}(X, W) = \rho = 0.3$, constructed with the VITA
(vine-to-anything) algorithm [@Gronneberg2022; @Foldnes2015]. Five exogenous
distributions were examined, each standardized to mean $0$ and variance $1$ so
that branches differ in shape rather than scale or location: normal, 
$X, W \sim \mathcal{N}(0, 1)$,
using a forced Gaussian copula. The remaining four combine non-normal margins 
with a non-Gaussian vine copula:
uniform, $t_5$, and two $\chi^2(2)$ conditions with $X$ and $W$ skewed in the 
same or in opposite directions. Given that, $M$ and $Y$ were generated from the 
structural equations above with independent Gaussian disturbances 
$\zeta_M \sim \mathcal{N}(0, \psi_M)$ and $\zeta_Y \sim \mathcal{N}(0, \psi_Y)$, 
and each latent variable's four 
indicators were formed as $z_{\eta j} = \lambda_j \eta + \delta_{\eta j}$, 
with loadings $\lambda = (1, 0.8, 0.7, 0.6)$ and Gaussian errors
$\delta_{\eta j} \sim \mathcal{N}(0, \theta_{\eta j})$ whose variances 
$\theta_{\eta j}$ were scaled to the target reliability $\omega \in \{0.5, 0.7\}$.

The nine (mis)specification conditions comprise one correctly specified condition
and four families of misspecification, each at two empirically calibrated 
magnitudes (standardized $0.25$ and $0.50$, detailed in Appendix A). 
Two are structural: an omitted direct effect of the moderator on the outcome 
(`c2`) and an omitted moderated direct effect (`c3`). The others are in the 
measurement model: a cross-loading of a mediator indicator on $X$ (`cl`) and a 
residual covariance between a mediator and an outcome indicator (`rcov`). 
In every case the effect is present in the 
population but omitted from the fitted model.

### Nonconvergence, Outliers, and Inadmissible Solutions

Following @Lonati2025, replications were screened for estimation failures and
inadmissible solutions before analysis. Because the compared estimators are
implemented across different packages, we applied an identical admissibility
check to all of them: a solution was flagged when any residual (unique) variance
was nonpositive (a Heywood case) or when the latent covariance matrix was not
positive definite. Each estimator was fit independently within a replication, so
a failure of one method produced a missing row for that method only, with the
captured error or warning retained for diagnostics. 

### Performance Measures

Performance was assessed with the measures implemented in the *simhelpers*
package [@Joshi2025], following recommendations in @Morris2019. For the index of
moderated mediation and the structural coefficients we report absolute and
relative bias, RMSE, the mean standard error and the SE/SD ratio,
confidence-interval coverage, and—where applicable—the rejection rate
(Type I error when $a_3 = 0$, power otherwise). Monte Carlo uncertainty in each
performance measure is quantified with jackknife Monte Carlo standard errors
[@Joshi2025].

Inference for the index of moderated mediation accounts for the skewness of the
product $a_3 b$: for $r = 1, \dots, R$ (with $R = 20{,}000$) we drew

$$
\bigl(a_3^{(r)}, b^{(r)}\bigr) \sim
\mathcal{N}_2\!\bigl((\hat a_3, \hat b),\, \hat\Sigma_{a_3 b}\bigr),
\qquad
\mathrm{imm}^{(r)} = a_3^{(r)}\, b^{(r)},
$$

where $\hat\Sigma_{a_3 b}$ is the estimated covariance of $(\hat a_3, \hat b)$,
and took the $2.5$th and $97.5$th empirical percentiles of
$\{\mathrm{imm}^{(r)}\}_{r=1}^{R}$ as the interval. The structural coefficients
and loadings use Wald (?) intervals from the model-implied parameter covariance
matrix. The same
implementation was applied to all three estimators for comparability.

### Reproducibility and Computational Details

All analyses were conducted in R [@RCoreTeam] using the *lavaan* [@Rosseel2012]
and *modsem* [@Slupphaug2025] packages for estimation, *covsim* [@Gronneberg2022]
and *rvinecopulib* [@Nagler2025] for data generation, and *simhelpers*
[@Joshi2025] for performance evaluation. Conditions were run in parallel with R's
*parallel* package; reproducible and independent random-number streams across
workers were obtained with the L'Ecuyer-CMRG generator seeded from a single
master seed. The computational environment was pinned with Nix via the *rix*
package [@rix] to ensure reproducibility of the computational environment, 
and the article was written with the *apaquarto* extension 
[@apaquarto]. All code and materials are available at the project repository.
